---
title: "2008 financial crisis in US and its impact on median income and housing prices."
author: "Studentnames and studentnumbers here Marine Latour, Bryan Berlijn, Toine van Bakel, Tyfarah Wedervoort, Shruti Ramasamy, Naomi van Leuvensteijn, Rocco vitale
"
date: "`r Sys.Date()`"
output: pdf_document
---

# Set-up your environment

```{r package_install, include=FALSE}
install.packages(c("ggplot2","tidyverse", "dplyr", "sf", "tigris", "viridis"))
```

```{r packages}
library(tidyverse)
library(readr)
library(dplyr)
library(ggplot2)
library(viridis)
library(tigris)
library(sf)
```

# Title Page

Include your names = "Marine Latour, Bryan Berlijn, Toine van Bakel, Tyfarah Wedervoort, Shruti Ramasamy, Naomi van Leuvensteijn, Rocco vitale"

Include the tutorial group number

Include your tutorial lecturer's name

# Part 1: Problem Motivation & Literature

## 1.1: Why was the 2008 Financial Crisis in the US a Social Problem?

The 2008 Financial Crisis, commonly referred to as the Great Recession, originated in the U.S.  and lasted from late 2007 to mid-2009. It was a major social problem due to its severe effects on median household income and housing prices, which significantly undermined economic security for millions of Americans. The crisis was mainly caused by excessive household leverage, particularly within the subprime mortgage sector, which triggered a steep decline in housing prices and a wave of mortgage defaults (Mian & Sufi, 2010a). The housing market collapse diminished household wealth and reduced consumer spending, intensifying the recession’s severity and duration. 

Mian and Sufi (2010a) demonstrate that households with higher leverage before the crisis faced substantially greater financial distress, leading to sharp reductions in durable consumption and higher unemployment rates at the county level. Furthermore, Morrissey et al. (2020) show that during the Great Recession, households with lower education levels, children, or without homeownership experienced significantly higher income and employment instability, an effect closely linked to increased financial hardship (pp. 3–4). Together, these economic and social disruptions illustrate why the Great Recession was not only a financial crisis but also a profound social crisis, affecting housing security, income stability, and overall well-being. 

## 1.2: What sources have identified this topic as a social problem?

Several peer-reviewed studies have recognized the 2008 Financial Crisis as a critical social problem. Mian and Sufi’s research (2010a, 2010b) emphasizes the role of household leverage and its consequences on the broader economy and local communities, highlighting the direct connection between financial market failures and deteriorating social conditions such as unemployment and poverty. Morrissey et al. (2020) extend this understanding by detailing how economic instability disproportionately affected vulnerable groups, including less-educated households and renters, thereby exacerbating social inequality during the recession. 

## 1.3: Has a part of this topic not been researched yet?

While extensive research has analyzed household leverage, foreclosures, unemployment, and macroeconomic impacts of the 2008 financial crisis such as by Mian & Sufi, 2010a, 2010b, Morrissey et al., 2020), less academic focus has been placed on the geographic variation in how the relationship between housing price shifts and median income evolved across states during the crisis. What remained unexplored is quantitatively assessing how differently these two indicators interacted on a state-by-state basis, thus creating regional disparities. 

## 1.4: How does the report bring new perspective on the topic?

This report creates a merged, state-level dataset to analyze income and housing prices together. Also, the following analysis will highlight where there was a misalignment between housing prices and income levels. Moreover, this report will visualize the regional variation among states in housing and income trends using plots and correlations. Furthermore, this report examines these patterns through the wealth effect. The wealth effect demonstrates how changes in prices of assets, especially property values, influences consumer spending. During the 2008 crisis, as house prices were declining, the confidence of consumers in the value of their property also fell. Thus, they perceived a reduction in household wealth. This negative wealth effect, in turn, reduced consumer spending, which consequently contributed to higher unemployment at the country level.  

# Part 2 - Data Sourcing

## 2.1 Load in the data

Preferably from a URL, but if not, make sure to download the data and store it in a shared location that you can load the data in from. Do not store the data in a folder you include in the Github repository!

```{r loading_data}
dataset <- midwest
```

midwest is an example dataset included in the tidyverse package

## 2.2 Provide a short summary of the dataset(s)

```{r}
head(dataset)
```

In this case we see 28 variables, but we miss some information on what units they are in. We also don't know anything about the year/moment in which this data has been captured.

These are things that are usually included in the metadata of the dataset. For your project, you need to provide us with the information from your metadata that we need to understand your dataset of choice.

## 2.3 Describe the type of variables included

Think of things like:

-   Do the variables contain health information or SES information?

-   Have they been measured by interviewing individuals or is the data coming from administrative sources?

*For the sake of this example, I will continue with the assignment...*

# Part 3 - Quantifying

## 3.1 Data cleaning

Say we want to include only larger distances (above 2) in our dataset, we can filter for this.

```{r data_cleaning}
mean(dataset$percollege)
```

Please use a separate 'R block' of code for each type of cleaning. So, e.g. one for missing values, a new one for removing unnecessary variables etc.

## 3.2 Generate necessary variables

Variable 1

```{r gen_var1}

```

Variable 2

```{r gen_var2}

```

## 3.3 Visualize temporal variation

```{r}

```

## 3.4 Visualize spatial variation

```{r visualise_map}

```

Here you provide a description of why the plot above is relevant to your specific social problem.

## 3.5 Visualize sub-population variation

What is the poverty rate by state?

```{r visualise_subpopulations}
dataset$inmetro <- dataset$inmetro %>% as.factor()
# Boxplot of poverty rate by state using the 'midwest' dataset
ggplot(dataset, aes(x = inmetro, y = percadultpoverty)) +
  geom_boxplot() +
  labs(
    title = "Distribution of Poverty Rates by Metropolitan status (Midwest counties)",
    x = "Metropolitan Area",
    y = "Poverty Rate of Adults (%)"
  ) +
  theme_minimal() +
  theme(
    legend.position = "right"
  )
```

Here you provide a description of why the plot above is relevant to your specific social problem.

## 3.6 Event analysis

Analyze the relationship between two variables.

```{r analysis}

```

Here you provide a description of why the plot above is relevant to your specific social problem.

# Part 4 - Discussion

## 4.1 Discuss your findings

# Part 5 - Reproducibility

## 5.1 Github repository link

Provide the link to your PUBLIC repository here: ...

## 5.2 Reference list

Use APA referencing throughout your document.
